---
title: "Dose-dependent interaction on a real network: sacubitril and valsartan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dose-dependent interaction on a real network: sacubitril and valsartan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_stan <- requireNamespace("cmdstanr", quietly = TRUE) ||
            requireNamespace("rstan", quietly = TRUE)
```

This vignette reproduces the sacubitril and valsartan blood-pressure analysis
from the methods paper. The network combines a valsartan dose-ranging axis, a
fixed-ratio combination diagonal, and the RATIO trial's off-diagonal arms, which
vary free sacubitril on a fixed valsartan backbone. The off-diagonal arms are
what make the interaction estimable.

## The data

Each row is one trial arm. The two component dose columns, `d_sac` and `d_val`,
carry the sacubitril and valsartan-equivalent doses, with zero meaning the
component is absent.

```{r data}
library(cdtmbnma)
sv <- sacval_example()
head(sv)
```

## Building the design

`cdt_data()` turns the arm-level table into the model design. The all-zero arm
in each study becomes the baseline, and the two components co-occur in the
combination arms, so one interaction pair is created.

```{r design}
d <- cdt_data(sv, study = "study", components = c("d_sac", "d_val"),
              outcome = "continuous", y = "y", se = "se",
              dstar = c(d_sac = 200, d_val = 320))
d
```

## Fitting

`cdt_fit()` compiles the Stan model on first use and samples. The bilinear
surface is the default. The reference doses passed as `dstar` set the dose
normalisation, so the interaction parameter reads as the effect at sacubitril
200 mg with valsartan 320 mg.

```{r fit, eval = run_stan}
fit <- cdt_fit(d, interaction = "bilinear", chains = 4,
               iter_warmup = 1000, iter_sampling = 1000, seed = 1)
summary(fit)
```

The component dose-response curves recover the expected shapes, valsartan
saturating early and the sacubitril add-on rising across the observed range.

```{r plot, eval = run_stan, fig.width = 7, fig.height = 3.2}
plot(fit)
```

## Prediction at an unobserved combination

`predict()` returns the relative effect against placebo at any dose
combination, including pairs that no trial observed. Here is the combination of
sacubitril 100 mg with valsartan 160 mg, which the diagonal and off-diagonal
arms bracket but never hit directly.

```{r predict, eval = run_stan}
predict(fit, newdata = data.frame(d_sac = 100, d_val = 160))
```

## Reading the interaction

A positive interaction on this outcome scale means sub-additivity, since lower
blood pressure is the benefit and the outcome is the change from baseline. The
paper reports the estimate as directionally sub-additive but not conclusive on
the off-diagonal evidence, and cautions that pooling fixed-ratio diagonal trials
inflates the apparent precision. Refitting with `interaction = "none"` gives the
additive benchmark, and comparing the two with `loo` on the stored `log_lik`
quantifies what the interaction buys.

```{r additive, eval = FALSE}
fit0 <- cdt_fit(d, interaction = "none")
# loo::loo_compare(loo::loo(fit$fit$draws("log_lik")),
#                  loo::loo(fit0$fit$draws("log_lik")))
```
